The most comprehensive treatment of ILL is in Gavin Bierman's thesis
\cite{Bierman:1994}.  There one finds the Proof Theory (Chapter 2),
i..e, the sequent calculus with cut-elimination, natural deduction and
axiomatic versions of ILL. Then (Chapter 3) a term assignment to the
natural deduction and to the sequent calculus versions are presented
with $\beta$-reductions and commutative conversions, and strong
normalization and confluence are proved for the resulting calculus. A
painstaking analysis of the rules of the labeled calculus leads to
the construction of a categorical model of ILL, in particular of the
exponential part, a main contribution of Bierman and of the Cambridge
school of the 1990s with respect to previous models by Seely and
Lafont.  Bellin \cite{Bellin:2014} presents a categorical model of
co-intuitionistic linear logic based on a dualization of Bierman
\cite{Bierman:1994} construction for ILL..

% <<<<<<< HEAD
The most comprehensive treatment of ILL is in Gavin Bierman's thesis  \cite{Bierman:1994}.
There one finds the Proof Theory (Chapter 2), i.e, the sequent calculus with cut-eliminaton, natural deduction and
axiomatic versions of ILL. Then (Chapter 3) a term assignment to the natural deduction and to the sequent calculus
versions are presented with $\beta$-reductions and commutative conversions, and strong normalization and confluence
are proved for the resulting calculus. A painstaking analysis of the rules of the labeled calculus leads to the construction
of a categorical model of ILL, in particular of the exponential part, a main contribution of Bierman and of the
Cambridge school of the 1990s with respect to previous models by Seely and Lafont.
Bellin \cite{Bellin:2012} presents a categorical model of co-intuitionistic linear logic based on a dualization of
Bierman \cite{Bierman:1994} construction for ILL..

Benton's work \cite{Benton:1994} on LNL logic, which we read in a TeX report, presents the categorical model 
for Linear-Non-Linear Intuitionistic logic LNL. Chapter 2 shows
how to obtain a LNL model from a Linear Category and viceversa. Versions of the sequent calculus for LNL
are considered  and cut-elimination is proved for one calculus. Then Natural Deduction is given with term
assignment and the categorical interpretation of a fragment of the natural deduction system. $\beta$ reductions
and commuting conversions are presented.
The present work follows Benton's paper aiming at a (non-trivial) dualization of it.

Bi-intuitionistic logic was introduced by C.Rauszer \cite{Rauszer:1974},  \cite{Rauszer:1974a}, \cite{Rauszer:1977} with
an algebraic and Kripke semantics and  \cite{Rauszer:1974a} a Gentzen style sequent calculus. Co-intuitionistic logic requires a multiple conclusion system, because of the cotensor in the linear case and of contraction right in the non-linear one. This raises the problem of the relations between intuitionistic implication and disjunction, and, dually, 
between subtraction and conjunction. Maheara and Deagalin's restriction  
 that only one formula occures in the succedent [antecedent] of the premise of an implication right [subtraction left] yields a calculus that does not satisfies cut-elimination already in the logic FILL that extends ILL with the cotensor \emph{par} 
as noticed by Schellinx \cite{Schellinx:1991}.

Benton's work \cite{Benton:1994} on LNL logic, which we read in a TeX
report, presents the categorical model for Linear-Non-Linear
Intuitionistic logic LNL. Chapter 2 shows how to obtain a LNL model
from a Linear Category and vice versa. Versions of the sequent calculus
for LNL are considered and cut-elimination is proved for one
calculus. Then Natural Deduction is given with term assignment and the
categorical interpretation of a fragment of the natural deduction
system. $\beta$ reductions and commuting conversions are presented.
The present work follows Benton's paper aiming at a (non-trivial)
dualization of it.

Bi-intuitionistic logic was introduced by C.Rauszer
\cite{Rauszer:1974}, \cite{Rauszer:1974a}, \cite{Rauszer:1977} with an
algebraic and Kripke semantics and \cite{Rauszer:1974a} a Gentzen
style sequent calculus. Co-intuitionistic logic requires a multiple
conclusion system, because of the tìco-tensor in the linear case and
of contraction right in the non-linear one. This raises the problem of
the relations between intuitionistic implication and disjunction, and,
dually, between subtraction and conjunction. Maheara and Deagalin's
restriction that only one formula occurs in the succedent
[antecedent] of the premise of an implication right [subtraction left]
yields a calculus that does not satisfies cut-elimination already in
the logic FILL that extends ILL with the cotensor \emph{par} as
noticed by Schellinx \cite{Schellinx:1991}.
>>>>>>> 9fbc5b8c7473d9b3d8c55af93a8f19afad83c78d
\begin{center}
\begin{tabular}{ccc}
\AxiomC{$\Gamma, A \vdash B$}
\RightLabel{$\limp$ R}
\UnaryInfC{$\Gamma \vdash A \limp B$}
\DisplayProof & \hskip1in\strut& 
\AxiomC{$ A \vdash B, \Delta$}
\RightLabel{$\lsub$ E}
\UnaryInfC{$A \lsub B \vdash \Delta$}
\DisplayProof 
\end{tabular}
\end{center}
As a counterexample, consider the sequent $p \Rightarrow q, r
\rightarrow ((p - q) \wedge r)$ given by Pinto and Uusatlu around 2003
\cite{Pinto-Uustalu:2010}, which is provable with cut but not cut-free
with Dragalin's restrictions.

Hyland and De Paiva introduced a sequent calculus for FILL labeled
with terms
\begin{center}
\begin{tabular}{c}
\AxiomC{$\overline{y}:\Gamma, x:A \vdash t:B, \overline{u}:\Delta$}
\RightLabel{$\limp$ R}
\UnaryInfC{$\overline{y}: \Gamma \vdash \lambda x:T A \limp B, \overline{u}:\Delta$}
\DisplayProof
\end{tabular} 
\end{center}
% <<<<<<< HEAD
where $x: A$ occurs in $t:B$ if and only if there is an ``essential dependency'' of $B$ from $A$. 
The restriction on the $\limp$ I is that $x$ does not occur in the terms  $\overline{u}:\Delta$.
The original term assignment did not guarantee cut-elimination, as noticed by Bierman \cite{Bierman:1996}; 
the assignment to \emph{par left} ($\oplus$ L)  had to be fine tuned, as indicated by Bellin \cite{Bellin:1997} 
\footnote{For \emph{par} left ($\oplus L$) we need the following, :
where $x: A$ occurs in $t:B$ if and only if there is an ``essential
dependency'' of $B$ from $A$.  The restriction on the $\limp$ I is
that $x$ does not occur in the terms $\overline{u}:\Delta$.  The
original term assignment did no guarantee cut-elimination, as noticed
by Bierman \cite{Bierman:1996}; the assignment to \emph{par left}
($\oplus$ L) had to be fine tuned, as indicated by Bellin
\cite{Bellin:1997}
\footnote{For \emph{par} left ($\oplus L$) we need the following:
% >>>>>>> 9fbc5b8c7473d9b3d8c55af93a8f19afad83c78d
\[
\AxiomC{$x:A, \overline{x}:\Gamma\vdash \overline{r}: \Delta\qquad y:B, \overline{y}:\Pi\vdash \overline{s}: \Lambda$}
\UnaryInfC{$z: A \oplus B, \overline{x}:\Gamma, \overline{y}:\Pi\vdash  \overline{r'}: \Delta, 
\overline{s'}: \Lambda $}
\DisplayProof
\]
where for $r'_i\in \overline{r'}$ and $s'_j \in \overline{s'}$ we have 
\[
\begin{tabular}{rll}
$r'_i =$ & $\mathtt{let}\ z\ \mathtt{be}\ x-\ \mathtt{in}\ r_i, \qquad$ & if $x$ occurs in $r_i$, \\
             & $r_i$                                                                              & otherwise.\\
$s'_j =$ & $\mathtt{let}\ z\ \mathtt{be}\ -y\ \mathtt{in}\ s_j, \qquad$ & if $y$ occurs in $s_j$, \\
             & $s_j$                                                                              & otherwise.\\
\end{tabular}
\]
We may introduce non-existent dependencies if we define always $r'_i =
\mathtt{let}\ z\ \mathtt{be}\ x-\ \mathtt{in}\ r_i.$
}

A detailed presentation of the term calculus for FILL with a full proof of cut elimination, is in \cite{EadesDePaiva2016}, 
where the correctness for a categorical semantics for FILL is also proved.
Another correct formalization of FILL, a sequent calculus with a relational annotation, was given by Bra\"uner and De Paiva \cite{BraunedDePaiva:1997}, with a proof of cut-elimination.
The second author \cite{Bellin:1997} gave a system of proof nets for FILL which sequentialize in the sequent calculus
with term assignment; the essential fact here is that $x:A$ occurs in $t:B$ if and only if there is a ``directed chain''
between $A$ and $B$ in the proof structure. Here cut elimination is proved by reduction to cut-elimination for proof nets.  

A detailed presentation of the term calculus for FILL with a full
proof of cut elimination, is in \cite{EadesDePaiva2016}, where the
correctness for a categorical semantics for FILL is also proved.
Another correct formalization of FILL, a sequent calculus with a
relational annotation, was given by Bra\"uner and De Paiva
\cite{BraunedDePaiva:1997}, with a proof of cut-elimination.  The
second author \cite{Bellin1997} gave a system of proof nets for FILL
which sequentialize in the sequent calculus with term assignment; the
essential fact here is that $x:A$ occurs in $t:B$ if and only if there
is a ``directed chain'' between $A$ and $B$ in the proof
structure. Here cut elimination is proved by reduction to
cut-elimination for proof nets.


A system of two-sided proof nets (in the style of natural deduction)
was given by Cockett and Seely \cite{Cockett:1997}.  For
Bi-Intuitionistic Linear Logic, they gave also a system of proof nets,
corresponding to a sequent calculus without annotations and
restrictions that therefore collapses into classical MLL.  Recently,
Clouston, Dawson, Goré and Tiu cite{Clouston} gave an annotation-free
formalization for BILL, alternative to sequent calculi, in the form of
deep-inference and display calculi for BILL. This calculus enjoys
cut-elimination and is relevant to the categorical semantics
bi-intuitionistic linear logic.
 

Tristan Crolard \cite{Crolard:2001,Crolard:2004} made an in-depth study of Rauszer's logic. In \cite{Crolard:2001} he showed that models of Rauszer logic (called ``subtractive logic'') based on bi-cartesian closed categories (with co-exponents) collapse to preorders.
He also studied models of subtractive logic and showed that its first order theory is constant-domain logic, thus it is not 
a conservative extension of intuitionistic logic.

Crolard \cite{Crolard:2004} develops the type theory for subtractive logic, extending a system of multiple conclusion
classical natural deduction with a connective of subtraction and then decorating proofs with a system of \cite{annotations
of dependencies} that allow to identify ``constructive proofs'': these are derivations where only the premise of an implication introduction depends on the discharged assumption and only the premise of a subtraction elimination depends on the discharged conclusion. Therefore Crolard's sequent calculus with annotations is not affected by the 
counterexamples to cut-eliminations. 

The type theory is Parigot $\lambda\mu$-calculus extended with operators for sums, products and subtraction. 
where the operators for subtraction introduction and elimination are understood as a calculus of co-routines.
A constructive system of co-routines is then obtained by imposing restrictions on terms corresponding to the restrictions
 on constructive proofs.  Annotation-free formalizations use the display calculus \cite{Gore:2000}, nested sequents
\cite{GorePostnieceTiu:2008} and deep inference \cite{Postniece:2009}.

Tristan Crolard \cite{Crolard:2001,Crolard:2004} made an in-depth
study of Rauszer's logic. In \cite{Crolard:2001} he showed that models
of Rauszer logic (called ``subtractive logic'') based on bi-cartesian
closed categories (with co-exponents) collapse to prerders.  He than
studied models of subtractive logic and showed that its first order
theory is constant-domain logic, thus it is not a conservative
extension of intuitionistic logic.  Crolard \cite{Crolard:2004}
develops the type theory for subtractive logic, extending a system of
multiple conclusion classical natural deduction with a connective of
subtraction and then decorating proofs with a system of
\cite{annotationsofdependencies} that allow to identify ``constructive
proofs'': these are derivations where only the premise of an
implication introduction depends on the discharged assumption and only
the premise of a subtraction elimination depends on the discharged
conclusion. Therefore Crolard's sequent calculus with annotations is
not affected by the counterexamples to cut-eliminations.

The type theory is Parigot $\lambda\mu$-calculus extended with
operators for sums, products and subtraction.  where the operators for
subtraction introduction and elimination are understood as a calculus
of co-routines.  A constructive system of co-routines is then obtained
by imposing restrictions on terms corresponding to the restrictions on
constructive proofs.  Annotation-free formalizations use the display
calculus \cite{Gore:2000}, nested sequents \cite{GorePostnieceTiu:2008}
and deep inference \cite{Postniece:2009}.

In a series of papers the second author gave a ``pragmatic''
interpretation interpretation of bi-intuitionism, where intuitionistic
and co-intuitionistic logic are interpreted as logics of the acts of
assertion and making a hypothesis, respectively, the interactions
between the two sides depending on negations (see, e.g,
\cite{Bellin:2014} Assertions, hypotheses, conjectures, expectations:
Rough-sets semantics and proof-theory.)  Here a term assignment for
co-intuitionistic logic has been derived from Crolard's definition but
independently from the $\lambda\mu$ framework. This calculus was used
here as a term assignment of Dual LNL logic.


To achieve the project outlined in the introduction of putting
together intuitionistic and co-intuitionistic adjoint logic in the
environment of BILL the definition of a suitable syntax for BILL will
play a role.


% More generally, notice also that co-intuitionistic logic also arises as a logic of refutations of intuitionistic sequents,
% dual to an intuitionistic logic of proofs \cite{Tranchini}.
% However, co-intuitionistic logic could also be regarded as \emph{opposite} to intuitionistic logic, rather than its dual.
% =======
% More generally, notice also that co-intuitionistic logic also arises
% as a logic of refutations of intuitionistic sequents, dual to an
% intuitionistic logic of proofs \cite{Tranchini}.  However,
% co-intuitionistic logic could also be regarded as \emph{opposite} to
% intuitionistic logic, rather than its dual.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main.tex
%%% End: 
% >>>>>>> 9fbc5b8c7473d9b3d8c55af93a8f19afad83c78d
